### 2022-10-04 Paper.tex

\documentclass{article}
\usepackage[utf8]{inputenc}
\title{\vspace{-2.75cm}634\vspace{-0.5cm}}
\author{Nicholas Ray}
\date{\vspace{-0.30cm}October 31 2022\vspace{-1cm}}
\usepackage[margin=1in]{geometry}
\usepackage{mathtools,amssymb,amsthm}
\usepackage{setspace}
\doublespacing
\usepackage[backend=biber, style=authoryear, maxbibnames=99, uniquelist=false]{biblatex}
\renewbibmacro{in:}{}
\renewbibmacro*{volume+number+eid}{%
  \printfield{volume}
  \setunit*{\addnbthinspace}
  \printfield{number}
  \setunit{\addcomma\space}
  \printfield{eid}}
\DeclareFieldFormat[article]{number}{\mkbibparens{#1}}
\AtEveryBibitem{
  \clearfield{issn}
  \clearfield{month}
  \clearfield{urlyear}
  \clearlist{language}
  \clearfield{note}
  \ifentrytype{online}{}{
    \clearfield{url}
  }
}
\addbibresource{ForeignAidBib.bib}
\begin{document}
\maketitle
\section*{Introduction}
%Look at Dreher's book, Kidd's international formal theory book, Hernadez 2017 paper, etc.
BRI (\cite{dreher2022}), Globalization (papers from week 6)?

Lots of academic attention on supply side provision of aid given China's entrance (e.g., \cite{dreher2018}, \cite{dreher2015}), but few explorations of the demand side after China has become a prevalent donor. Also research on how Chinese aid affects recipients in various ways (e.g., \cite{martorano2020}, \cite{bader2015}).
This paper seeks to address this shortage in the literature, aiming to understand how recipient countries choose between the vast number of credible donors. Relevant literature: \cite{kilama2016a}, \cite{hernandez2017}, \cite{li2017a}, \cite{isaksson2018}, \cite{humphrey2019}, \cite{broich2017a}.

Research Question. Given the literature, the decision-making of recipient countries no longer seems so trivial (i.e., why are countries choosing to receive aid money from China?). The literature implies that there should be a healthy mix of aid donor money given the effects shown. Some strategy is no involved on behalf of the recipient country; if we are not seeing a diversified diet in aid receipt, why not? Seems best.

\section*{Theory}
Explanation

\section*{Methodology}
Data \& Variables

Research Design

\pagebreak
\printbibliography
\end{document}