### 2022-12-08 Other/Future.tex
Updates from Overleaf

Weak logical link between Western entities giving fewer concessions and the conclusion that recipients should diversify borrowing

We're implicitly assuming that loans are sufficiently easy for African countries to obtain from both China and West. Big problem