### 2022-12-08 Paper.tex

\documentclass{article}
\usepackage[utf8]{inputenc}
\title{\vspace{-2.75cm}On Government Borrowing Behavior\vspace{-0.5cm}}
\author{Nicholas Ray}
\date{\vspace{-0.30cm}December 7 2022\vspace{-1cm}}
\usepackage[margin=1in]{geometry}
\usepackage{mathtools,amssymb,amsthm}
\usepackage{setspace}
\usepackage{tikz}
\usepackage{sgame}
\doublespacing
\usepackage[backend=biber, style=authoryear, maxbibnames=99, uniquelist=false]{biblatex}
\renewbibmacro{in:}{}
\renewbibmacro*{volume+number+eid}{%
  \printfield{volume}
  \setunit*{\addnbthinspace}
  \printfield{number}
  \setunit{\addcomma\space}
  \printfield{eid}}
\DeclareFieldFormat[article]{number}{\mkbibparens{#1}}
\AtEveryBibitem{
  \clearfield{issn}
  \clearfield{month}
  \clearfield{urlyear}
  \clearlist{language}
  \clearfield{note}
  \ifentrytype{online}{}{
    \clearfield{url}
  }
}
\addbibresource{ForeignAidBib.bib}
\begin{document}
\maketitle
\section*{Introduction}
Since engaging in a ``going out'' strategy around the turn of the century, the People's Republic of China (China) has successfully become the ``lender of first resort'' for developing countries \parencite[1]{dreher2022}. There has been an abundance of scholarly work attempting to understand the potential effects of China's contemporary prominence in the global finance arena. Much of this literature focuses on how the international financing behavior of more traditional, Western donors have been affected (e.g., \cite{humphrey2019}; \cite{kilama2016a}) or how the effectiveness of their money has changed (e.g., \cite{blair2022}; \cite{gehring2022}).

Little of this work has looked at how the behavior of recipients of international finance has updated since the availability of Chinese money. This lack of attention is despite recent work that implies that recipient countries may extract benefits from their loan choices in a more Chinese global finance environment. While the literature implies that recipients stand to gain from receiving both Chinese and Western finance and the subsequently more competitive terms, observations of borrowing practices in Africa do not seemingly reflect that countries take relatively similar levels of both loan types.

I argue that this is due to time-inconsistency problems where countries prefer Western or Chinese loans over the other in the short-run. Understanding the debt composition of developing countries is an increasingly important issue as public debt continues to rise. This research could also shed light on literature concerned about China's interaction with the liberal international order (LIO) and spark a new research agenda studying recipient behavior.

\section*{What Demands Explanation and Why}
There is evidence that countries are offered fewer loan conditions from the World Bank (\cite{hernandez2017}) and that Western aid is less sensitive to policy and institutional quality (\cite{annen2021}) when they are recipients of Chinese finance. Although Dreher et al. (2021)\nocite{dreher2021} did not find a sizable negative effect of Chinese finance on Western aid's effectiveness, they did find that aid from the United States (U.S.) seems to be more effective in the absence of Chinese finance. In sum, this evidence suggests that receiving Chinese loans may induce the additional benefit of better financial terms from subsequent Western lenders.

There is also evidence that Chinese finance increases with Western development assistance, which Dreher et al. (2018)\nocite{dreher2018} interpret as competitive behavior from China. Taken together, the literature seems to imply that there is some financial competition between the West and China that recipients may be able to take advantage of to procure more favorable borrowing terms. It follows that recipients should have a relatively diversified composition of debt between Chinese and Western lenders to stimulate this competition and term improvement.

However, it is at least observed in Africa that the proportion of Chinese loans to Western loans varies substantially between counties and further that some countries almost entirely borrow from either China or the West when taking money from governments. For example, compare table 1 and table 2, which show the estimated composition of public debt in Cabo Verde and Djibouti, respectively (\cite{jubileedebtcampaign2018}). Cabo Verde owes only 1\% of its estimated total debt to China while debt to Portugal and other unspecified countries exceeds 25\%. In Djibouti, meanwhile, 68\% of its estimated debt can be attributed to China.

\begin{table}
    \centering
    \includegraphics[scale=0.5]{Figures/CaboVerde.png}
    \caption{Estimated Debt for Cabo Verde}
    \vspace{1cm}
    \includegraphics[scale=0.5025]{Figures/Djibouti.png}
    \caption{Estimated Debt for Djibouti}
\end{table}

Thus, the variance in debt composition and the seeming failure of countries to borrow in a more equal manner between China and Western entities must be explained. In other words, how do recipient countries decide between taking loans from China and the West? Taking loans is costly, at least in the long-run, and recipients presumably want to maximize the payoff they receive from borrowing. Related literature seems to not have fully considered recipient decision-making with regard to choosing a borrower, instead largely focusing on macroeconomic and political determinants of government debt (e.g., \cite{bittencourt2015} ; \cite{swamy2015}) or the determinants of lender behavior (e.g., \cite{dreher2018}).

\section*{My Explanation}
I argue that the variation observed in African borrowing practices is a result of strategic considerations on behalf of recipients. Specifically, I argue that some countries take Chinese and Western finance at substantially different rates due to dissimilarities in the conditions of Western and Chinese loans and the subsequent evaluations of loan value by recipients.

There is evidence that China might be simultaneously more strict on the economic conditions on their finance (e.g., \cite{dreher2018}) while also less strict on the political conditions imposed (e.g., \cite{dreher2018}). Western loans, by contrast, seem more politically strict than economic (\cite{hernandez2017}). This dissimilarity leads to time-inconsistency problems, where in the short-term countries prefer either less severe economic or political conditions instead of minimizing the joint conditionality over the long-term.

One implication is that more democratic countries should be more sensitive to explicitly political conditions and, thus, prefer Western loans $ceteris\;paribus$. This could also explain Dreher et al.'s (2018) findings that Chinese loans seem to flow to more corrupt countries and those with poorer political institutions.

\subsubsection*{Figure 1: Non-Democratic Country Loan Preference, $ceteris\;paribus$}
\begin{tikzpicture}[scale=2,declare function={F(\x) = 1.7*(2*\x-\x*\x)-0.2;}]
  \draw[-latex] (0,-0.2) -- (0,2) node[left]{$Utility$};
  \draw[-latex] (-0.2,0) -- (2,0) node[below right]{$Political\;Conditionality$};
  \draw[domain=0.1:1.9,samples=10,smooth,thick] plot (\x, {F(\x)}); 
  \coordinate (max) at (0,{F(1)});
  \draw[dashed] (max) -| (1,0) node[pos=1,below]{China};
  \coordinate (max) at (0,{F(1.5)});
  \draw[dashed] (max) -| (1.5,0) node[pos=1,below]{West};
\end{tikzpicture}

\section*{Research Design}
The implication above leads to the following hypotheses:
\begin{align*}
    H_{1a}&:\;\text{Democracies should receive more Western loans than Chinese loans}\\
    H_{1b}&:\;\text{Non-democracies should receive more Chinese loans than Western loans}\\
\end{align*}
To test these hypotheses I intend to focus on the international loan environment in Africa since from the years 2000 to 2020 due the continent being a major area of focus for China's finance projects and this being the time period most relevant for Chinese finance. I will fit a logit/probit model with the lender type (e.g., majority Western or not) as my dependent variable. The main explanatory variables will be regime type and a measure for loan conditionality. Ideally, an instrumental variable will be used for regime type since there is theoretically expected to be simultaneity present. Further, I will include gross domestic product (GDP) per capita as a control for recipient need and an additional control for the gross sum of the loan to ensure comparison.

I plan to use data from William \& Mary's ``Aiddata'' lab for Chinese loans (\cite{custer2021}) and data from the Organization for Economic Co-operation and Development (OECD) on Western loans (\cite{organizationforeconomicco-operationanddevelopment2022a}). GDP per capita stems from the United Nations (\cite{unitednationsstatisticsdivision2019}). For regime type, I intend to use an updated version of Cheibub et al.'s dictatorship and democracy index (\cite{cheibub2010}).

\nocite{curini2020}
\pagebreak
\printbibliography
\end{document}